\textmd{\bf Cryptography: Birthday Paradox}

### 2019-01-17 draft.tex
Fix minor bug

    \section{}
    \subsection{}
    Theorem 1.1. Let $S = \{1, 2, \cdots, N\}$. For n times, uniformly randomly draw one element from set S with replacement. Let $x_t$ be the element we draw at time t. Then $\forall p > 0$, there exists a constant $C_1$ such that when $n \geq C_1 \sqrt{N}$, we have
    $$Pr[\exists 1\leq i, j \leq n, i\not = j \text{ such that } x_i = x_j] > p$$.
    \begin{proof}
        Let X denote the event that $\exists i, j \leq t, i\not = j \text{ such that } x_i = x_j$, then we have
        \begin{equation}
            \begin{aligned}
                Pr[\bar{X}] &= \frac{N(N-1)\cdots (N-n+1)}{N^n}\\
                &= \prod_{i=1}^{n-1}(1 - \frac{i}{N})\\
                &\leq \prod_{i=1}^{n-1} exp(-\frac{i}{N})\\
                &= exp(-\frac{n(n-1)}{2N})
            \end{aligned}
        \end{equation}
        Let $C_1 = \sqrt{-2ln(1 - p)} + 1$. Then when $n \geq C_1\sqrt{N}$, we have
        \begin{equation}
            \begin{aligned}
                n(n-1) > (1 + \sqrt{-2ln(1 - p)\cdot N})(\sqrt{-2ln(1 - p)\cdot N}) > 2ln(2)\cdot N
            \end{aligned}
        \end{equation}
        Which is equivalent to $-\frac{n(n-1)}{2N} < ln(1 - p)$.
        Thus use (1) we have
        $$Pr[\bar{X}] \leq exp(-\frac{n(n-1)}{2N}) < 1 - p$$
        So we have
        $$Pr[X] > p$$
    \end{proof}
    \subsection{}
    Lemma 1.1. For positive integer $n < N$, we have
    $$\sum_{i=1}^{n-1}ln(1 - \frac{i}{N}) >-\frac{n^2}{N}$$
    \begin{proof}
        Notice that $\forall x\in [1 -\frac{i+1}{N}, 1 - \frac{i}{N}]$ $(0\leq i\leq n)$, we have $ln(x) < ln(1 - \frac{i}{N})$. Thus
        $$\frac{1}{N}ln(1 - \frac{i}{N}) \geq \int_{1 - \frac{i+1}{N}}^{1 - \frac{i}{N}}ln(x)dx$$
        Thus we have
        \begin{equation}
            \begin{aligned}
                \frac{1}{N}\sum_{i=1}^{n-1}ln(1 - \frac{i}{N}) &\geq \int_{1-\frac{n}{N}}^1 ln(x)dx\\
                &= (xln(x) - x)|_{1-\frac{n}{N}}^1\\
                &= -\frac{n}{N} - (1 - \frac{n}{N})ln(1 - \frac{n}{N})\\
                &> -\frac{n}{N} - (1 - \frac{n}{N})(-\frac{n}{N})\\
                &= - \frac{n^2}{N^2}
            \end{aligned}
        \end{equation}
        Thus
        $$\sum_{i=1}^{n-1}ln(1 - \frac{i}{N}) > -\frac{n^2}{N}$$
    \end{proof}
    Theorem 1.2. Let $S = \{1, 2, \cdots, N\}$. For n times, uniformly randomly draw one element from set S with replacement. Let $x_t$ be the element we draw at time t. Then $\forall p > 0$, there exists a constant $C_2$ such that when $n \leq C_2 \sqrt{N}$, we have
    $$Pr[\exists 1\leq i, j \leq n, i\not = j \text{ such that } x_i = x_j] < p$$.
    \begin{proof}
        Let X denote the event that $\exists i, j \leq t, i\not = j \text{ such that } x_i = x_j$.

@@

        Notice that
        \begin{equation}
            \begin{aligned}
                &\quad Pr_{D_t^n}(X_1\cup X_2 | Y = 2) - Pr_{D_{t+1}^n}(X_1\cup X_2 | Y = 2)\\
                &= \frac{1}{(D_t[1] + D_t[2])^2}(D_t[1]^2 + D_t[2]^2 - (\frac{1}{N})^2 - (D_t[1] + D_t[2] - \frac{1}{N})^2)\\
                &= -\frac{2}{(D_t[1] + D_t[2])^2}(D_t[1] - \frac{1}{N})(D_t[2] - \frac{1}{N})\\
                &> 0
            \end{aligned}
        \end{equation}
        And for any distribution D over S we have
        \begin{equation}
            Pr_{D^n}(X_1\cup X_2 | Y = i) =
            \begin{cases}
                0 & i = 0, 1\\
                1 & i \geq 3
            \end{cases}
        \end{equation}
        Thus
        $$\forall i, Pr_{D_t^n}(X_1\cup X_2 | Y = i) \geq Pr_{D_{t+1}^n}(X_1\cup X_2 | Y = i)$$
        Since we only adjust $D_t[0], D_t[1]$,
        $$Pr_{D_t^n}[\bigcup_{k = 3}^N X_k] = Pr_{D_{t+1}^n}[\bigcup_{k = 3}^N X_k]$$
        $$\forall i, Pr_{D_t^n}(Y = i | \bigcap_{k = 3}^N \bar{X_k}) = Pr_{D_{t+1}^n}(Y = i | \bigcap_{k = 3}^N \bar{X_k})$$
        So consider equation (6) and we get
        $$Pr_{D_t^n}[X] \geq Pr_{D_{t+1}^n}[x]$$
    \end{proof}

@@

    By induction, $\forall p$, we can find a constant $C_1$ such that
    $$Pr[X] > \frac{1+p}{2}$$
    for all N. Let $C_2$ be another constant such that $\frac{1+p}{2} \cdot (1 - (\frac{1}{e})^{C_2}) > p$ and $(1 - exp(-\frac{C_1 + C_2}{4})) > p$. We divide the whole drawing process into two steps:
    \begin{enumerate}
        \item First, draw $M_1 = C_1 \cdot N^{\frac{d-2}{d}} + C_2 \cdot N^{\frac{d-1}{d}}$ elements from S. Let set A be the set of all the elements that has been drawn for at least once.
        \item Second, draw $M_2 = 2(C_1 + C_2)\cdot N^{\frac{d-1}{d}}$ elements from S. Let $Y$ be the number of times that an element is drawn from set A.
    \end{enumerate}
    Now we consider 2 possible cases of the size of A.

    \subsubsection{First Case}
    If $|A| \leq n^{\frac{d-1}{d}}$:

    By assumption, after drawing for $C_1 \cdot N^{\frac{d-2}{d}}$ times, let event $E_1$ be that there exists an element $c_0$ in A that has been drawn for at least $d-1$ times, by assumption, given that $|A| \leq n^{\frac{d-1}{d}}$, we have $Pr(E_1) > \frac{p+1}{2}$.

@@

    Let $X_i$ be the indicater random variable of whether the i th draw in step 2 draws an element in A. In other words,
    \begin{equation}
        X_i = \begin{cases}
            1 & \text{if the i th draw draws an element from A} \\
            0 & \text{otherwise}
        \end{cases}
    \end{equation}
    Then we have $Y = \sum_{i=1}^{M_2}X_i$. And $X_i$s are i.i.d. Bornoulli random variables. So use Chernoff Bound
    $$ Pr(X < (1 - \delta)\mu) \leq e^{-\frac{\delta^2 \mu}{2}}$$
    We have
    \begin{equation}
        Pr[Y \leq \frac{1}{2}E[Y]] \leq e^{-\frac{E[Y]}{8}}
    \end{equation}
    Here 
    $$E[Y] = \frac{|A|}{N} \cdot 2(C_1 + C_2)\cdot N^{\frac{d-1}{d}} \geq 2\cdot (|A|)^{\frac{d-2}{d-1}}$$
    And
    $$E[Y] = \frac{|A|}{N} \cdot 2(C_1 + C_2)\cdot N^{\frac{d-1}{d}} \geq 2\cdot (C_1 + C_2)\cdot N^{\frac{d-2}{d}}$$
    Thus we have
    $$Pr[Y > (|A|)^{\frac{d-2}{d-1}}] \geq 1 - exp(-\frac{(C_1 + C_2)\cdot N^{\frac{d-2}{d}}}{4}) \geq 1 - exp(-\frac{C_1 + C_2}{4})$$
    Let the event $E_3$ be that there exists an elemnet $c_1$ in A such that $c_1$ has been drawn for at least $d-1$ times in step 2. By induction, we know in step 2, the conditional probability
    $$Pr[E_3 | Y > (|A|)^{\frac{d-2}{d-1}}] > \frac{1+p}{2}$$
    Thus we have
    \begin{equation}
        Pr[E_3] = Pr[E_3 | Y > (|A|)^{\frac{d-2}{d-1}}]\cdot Pr[Y > (|A|)^{\frac{d-2}{d-1}}] > \frac{p+1}{2}\cdot (1 - exp(-\frac{C_1 + C_2}{4})) > p
    \end{equation}
    Since the event $E_3$ gaurantees that $c_1$ has been drawn for at least d times (at least (d-1) in step 2, and at least 1 in step 1), we proved that
    $$Pr[X | |A| > N^{\frac{d-1}{d}}] > p$$

>>>>

    \section{}
    \subsection{}
    Theorem 1.1. Let $S = \{1, 2, \cdots, N\}$. For n times, uniformly randomly draw one element from set S with replacement. Let $x_t$ be the element we draw at time t. Then $\forall p > 0$, there exists a constant $C_1$ such that when $n \geq C_1 \sqrt{N}$, we have
    $$Pr[\exists 1\leq i, j \leq n, i\not = j \text{ such that } x_i = x_j] > p$$.
    \begin{proof}
        Let X denote the event that $\exists i, j \leq t, i\not = j \text{ such that } x_i = x_j$, then we have
        \begin{equation}
            \begin{aligned}
                Pr[\bar{X}] &= \frac{N(N-1)\cdots (N-n+1)}{N^n}\\
                &= \prod_{i=1}^{n-1}(1 - \frac{i}{N})\\
                &\leq \prod_{i=1}^{n-1} exp(-\frac{i}{N})\\
                &= exp(-\frac{n(n-1)}{2N})
            \end{aligned}
        \end{equation}
        Let $C_1 = \sqrt{-2ln(1 - p)} + 1$. Then when $n \geq C_1\sqrt{N}$, we have
        \begin{equation}
            \begin{aligned}
                n(n-1) > (1 + \sqrt{-2ln(1 - p)\cdot N})(\sqrt{-2ln(1 - p)\cdot N}) > -2ln(1-p)\cdot N
            \end{aligned}
        \end{equation}
        Which is equivalent to $-\frac{n(n-1)}{2N} < ln(1 - p)$.
        Thus use (1) we have
        $$Pr[\bar{X}] \leq exp(-\frac{n(n-1)}{2N}) < 1 - p$$
        So we have
        $$Pr[X] > p$$
    \end{proof}
    \subsection{}
    Lemma 1.1. For positive integer $n < N$, we have
    $$\sum_{i=1}^{n-1}ln(1 - \frac{i}{N}) >-\frac{n^2}{N}$$
    \begin{proof}
        Notice that $\forall x\in [1 -\frac{i+1}{N}, 1 - \frac{i}{N}]$ $(0\leq i\leq n)$, we have $ln(x) < ln(1 - \frac{i}{N})$. Thus
        $$\frac{1}{N}ln(1 - \frac{i}{N}) \geq \int_{1 - \frac{i+1}{N}}^{1 - \frac{i}{N}}ln(x)dx$$
        Thus we have
        \begin{equation}
            \begin{aligned}
                \frac{1}{N}\sum_{i=1}^{n-1}ln(1 - \frac{i}{N}) &\geq \int_{1-\frac{n}{N}}^1 ln(x)dx\\
                &= (xln(x) - x)|_{1-\frac{n}{N}}^1\\
                &= -\frac{n}{N} - (1 - \frac{n}{N})ln(1 - \frac{n}{N})\\
                &> -\frac{n}{N} - (1 - \frac{n}{N})(-\frac{n}{N})\\
                &= - \frac{n^2}{N^2}
            \end{aligned}
        \end{equation}
        Thus
        $$\sum_{i=1}^{n-1}ln(1 - \frac{i}{N}) > -\frac{n^2}{N}$$
    \end{proof}
    Theorem 1.2. Let $S = \{1, 2, \cdots, N\}$. For n times, uniformly randomly draw one element from set S with replacement. Let $x_t$ be the element we draw at time t. Then $\forall p > 0$, there exists a constant $C_2$ such that when $n \leq C_2 \sqrt{N}$, we have
    $$Pr[\exists 1\leq i, j \leq n, i\not = j \text{ such that } x_i = x_j] < p$$.
    \begin{proof}
        Let X denote the event that $\exists i, j \leq t, i\not = j \text{ such that } x_i = x_j$.

@@

        Notice that
        \begin{equation}
            \begin{aligned}
                &\quad Pr_{D_t^n}(X_1\cup X_2 | Y = 2) - Pr_{D_{t+1}^n}(X_1\cup X_2 | Y = 2)\\
                &= \frac{1}{(D_t[1] + D_t[2])^2}(D_t[1]^2 + D_t[2]^2 - (\frac{1}{N})^2 - (D_t[1] + D_t[2] - \frac{1}{N})^2)\\
                &= -\frac{2}{(D_t[1] + D_t[2])^2}(D_t[1] - \frac{1}{N})(D_t[2] - \frac{1}{N})\\
                &> 0
            \end{aligned}
        \end{equation}
        And for any distribution D over S we have
        \begin{equation}
            Pr_{D^n}(X_1\cup X_2 | Y = i) =
            \begin{cases}
                0 & i = 0, 1\\
                1 & i \geq 3
            \end{cases}
        \end{equation}
        Thus
        $$\forall i, Pr_{D_t^n}(X_1\cup X_2 | Y = i) \geq Pr_{D_{t+1}^n}(X_1\cup X_2 | Y = i)$$
        Since we only adjust $D_t[1], D_t[2]$,
        $$Pr_{D_t^n}[\bigcup_{k = 3}^N X_k] = Pr_{D_{t+1}^n}[\bigcup_{k = 3}^N X_k]$$
        $$\forall i, Pr_{D_t^n}(Y = i | \bigcap_{k = 3}^N \bar{X_k}) = Pr_{D_{t+1}^n}(Y = i | \bigcap_{k = 3}^N \bar{X_k})$$
        So consider equation (5), (6) and we get
        $$Pr_{D_t^n}[X] \geq Pr_{D_{t+1}^n}[x]$$
    \end{proof}

@@

    By induction, $\forall p$, we can find a constant $C_1$ (to make it more convenient, we do not use notation $C_1(\frac{p+1}{2}, d)$ here, but make sure constants should be independent with N) such that
    $$Pr[X] > \frac{1+p}{2}$$
    for all N. Let $C_2$ be another constant such that $\frac{1+p}{2} \cdot (1 - (\frac{1}{e})^{C_2}) > p$ and $(1 - exp(-\frac{C_1 + C_2}{4})) > p$. We divide the whole drawing process into two steps:
    \begin{enumerate}
        \item First, draw $M_1 = C_1 \cdot N^{\frac{d-2}{d}} + C_2 \cdot N^{\frac{d-1}{d}}$ elements from S. Let set A be the set of all the elements that has been drawn for at least once.
        \item Second, draw $M_2 = 2(C_1 + C_2)\cdot N^{\frac{d-1}{d}}$ elements from S. Let $Y$ be the number of times that an element is drawn from set A.
    \end{enumerate}
    Now we consider 2 possible cases of the size of A.

    \subsubsection{First Case}
    If $|A| \leq N^{\frac{d-1}{d}}$:

    By assumption, after drawing for $C_1 \cdot N^{\frac{d-2}{d}}$ times, let event $E_1$ be that there exists an element $c_0$ in A that has been drawn for at least $d-1$ times, by assumption, given that $|A| \leq N^{\frac{d-1}{d}}$, we have $Pr(E_1) > \frac{p+1}{2}$.

@@

    Let $X_i$ be the indicater random variable of whether the i th draw in step 2 draws an element in A. In other words,
    \begin{equation}
        X_i = \begin{cases}
            1 & \text{if the i th draw draws an element from A} \\
            0 & \text{otherwise}
        \end{cases}
    \end{equation}
    Then we have $Y = \sum_{i=1}^{M_2}X_i$. And $X_i$s are i.i.d. Bornoulli random variables. So use Chernoff Bound
    $$ Pr(X < (1 - \delta)\mu) \leq e^{-\frac{\delta^2 \mu}{2}}$$
    where $0\leq \delta \leq 1$, X is the sum of the random variables, $\mu$ is the expected value of the sum.
    We have
    \begin{equation}
        Pr[Y \leq \frac{1}{2}E[Y]] \leq e^{-\frac{E[Y]}{8}}
    \end{equation}
    Here 
    $$E[Y] = \frac{|A|}{N} \cdot 2(C_1 + C_2)\cdot N^{\frac{d-1}{d}} \geq 2\cdot (|A|)^{\frac{d-2}{d-1}}$$
    And
    $$E[Y] = \frac{|A|}{N} \cdot 2(C_1 + C_2)\cdot N^{\frac{d-1}{d}} \geq 2\cdot (C_1 + C_2)\cdot N^{\frac{d-2}{d}}$$
    Thus we have
    $$Pr[Y > (|A|)^{\frac{d-2}{d-1}}] \geq 1 - exp(-\frac{(C_1 + C_2)\cdot N^{\frac{d-2}{d}}}{4}) \geq 1 - exp(-\frac{C_1 + C_2}{4})$$
    Let the event $E_3$ be that there exists an elemnet $c_1$ in A such that $c_1$ has been drawn for at least $d-1$ times in step 2. By induction, we know in step 2, the conditional probability
    $$Pr[E_3 | Y > (|A|)^{\frac{d-2}{d-1}}] > \frac{1+p}{2}$$
    Thus we have
    \begin{equation}
        Pr[E_3] = Pr[E_3 | Y > (|A|)^{\frac{d-2}{d-1}}]\cdot Pr[Y > (|A|)^{\frac{d-2}{d-1}}] > \frac{p+1}{2}\cdot (1 - exp(-\frac{C_1 + C_2}{4})) > p
    \end{equation}
    Since the event $E_3$ gaurantees that $c_1$ has been drawn for at least d times (at least (d-1) in step 2, and at least 1 in step 1), we proved that
    $$Pr[X | |A| > N^{\frac{d-1}{d}}] > p$$
